---
title: "Hw 4"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DESeq2)
library(dplyr)
library(apeglm)
library(vsn)
library(pheatmap)
library(RColorBrewer)
```


# 1. Load Data
```{r}
data <- read.csv('GSE37704.csv')
rownames(data) <- data[,1]
data <- data[,-1]
```


# 2. Preprocess Data

## 2.1 Normalize Data by count per million
```{r}
normData <- lapply(data, function(x){return(x/sum(x) * 1000000)})
df <- as.data.frame(normData, row.names = rownames(data), col.names = names(normData))
summary.data.frame(df)
```

## 2.2 Log normalized data
```{r}
logDf <- log(df+1)
summary.data.frame(logDf)
```


## 2.3 cut data
```{r}
s <- df %>%
  dplyr::add_rownames() %>%
  dplyr::filter( (control_1 >= 5) | (control_2 >= 5) | (control_3 >= 5) 
         | (Hoxa1KN_1 >= 5) | (Hoxa1KN_2 >= 5) | (Hoxa1KN_3 >= 5)) %>%
  dplyr::select(rowname)
cuttedData <- data %>%
  dplyr::add_rownames() %>%
  dplyr::inner_join(s, by='rowname') 
cuttedData <- as.data.frame(cuttedData)
rownames(cuttedData) <- cuttedData[,1]
cuttedData <- cuttedData[,-1]
print("")
nrow(cuttedData)
summary(cuttedData)

```



# 3 Deseq 2
## 3.1 Generate result
```{r}

colna <- colnames(cuttedData)
coldata <- as.data.frame(c('control','control','control','Hoxa1KN','Hoxa1KN','Hoxa1KN'))
colnames(coldata) <- c('condition')
rownames(coldata) <- colna

head(coldata)


head(cuttedData)

dds <- DESeq2::DESeqDataSetFromMatrix(countData = as.matrix(cuttedData), colData = coldata, design = ~ condition )


dds

dds <- DESeq2::DESeq(dds)
res <- DESeq2::results(dds)
res

summary(res)
```

## 3.2 Cut result by FDR = 0.05 and abs(lfc) > 1
```{r}
res05 <- DESeq2::results(dds, alpha=0.05, lfcThreshold=1, altHypothesis = 'greaterAbs')
res05inc <- DESeq2::results(dds, alpha=0.05, lfcThreshold=1, altHypothesis = 'greater')
res05decs <- DESeq2::results(dds, alpha=0.05, lfcThreshold=1, altHypothesis = 'less')

summary(res05)
summary(res05inc)
summary(res05decs)
```



## 3.3 Get Gene List, sort by lfc
```{r}
UpGeneFilter <- which(  !is.na(res$padj) & 
                                      !is.na(res$log2FoldChange) & 
                                      res$padj < 0.05 & 
                                      res$log2FoldChange > 1)

UpGeneslfc <- res$log2FoldChange[UpGeneFilter]
UpGeneList <- rownames(res)[UpGeneFilter][order(-UpGeneslfc)]

DownGeneFilter <- which(  !is.na(res$padj) & 
                          !is.na(res$log2FoldChange) & 
                          res$padj < 0.05 & 
                          res$log2FoldChange < -1)
DownGeneslfc <- res$log2FoldChange[DownGeneFilter]
DownGeneList <- rownames(res)[DownGeneFilter][order(DownGeneslfc)]

AllGeneList <- rownames(res)


length(AllGeneList)
length(UpGeneList)
length(DownGeneList)

```

## 3.4 Refine hierarchical clustering
Use all differentially expressed gene
```{r}
vsd <- vst(dds, blind=FALSE)
ntd <- normTransform(dds)
resLFC <- DESeq2::lfcShrink(dds, coef="condition_Hoxa1KN_vs_control", type="apeglm")

select2 <- order(abs(resLFC$log2FoldChange[which(resLFC$padj<0.05&abs(resLFC$log2FoldChange)>1)]),
                decreasing=TRUE)

select_diff <- assay(ntd)[select2,]-rowMeans(assay(ntd)[select2,])

select3 <- order(rowMeans(select_diff[,1:3]), decreasing = TRUE)
#df <- as.data.frame(colData(dds)[,c("condition","type")])
index <- select2[select3]
length(index)
mySd <- sd( assay(ntd)[index,]-rowMeans(assay(ntd)[index,]) )
pheatmap(( assay(ntd)[index,]-rowMeans(assay(ntd)[index,]) )/mySd, cluster_rows=FALSE, show_rownames=FALSE,
         cluster_cols=TRUE)

```




```{r}
library(biomaRt)
library(ALL)
library(gage)
library("org.Hs.eg.db")
library(pathview)
library(png)
library(GO.db)
library(GOstats)
```


# 4. Enrichment: TopGo Solution
I failed to use topGo package. Instead, I use GOStats package.
## 4.1 Load Ensmbl to Go_id table 
```{r, eval=FALSE}
if(!file.exists('EG2GO.RDS')){
  grch37 = useEnsembl(biomart="ensembl", dataset = "hsapiens_gene_ensembl")
  EG2GO <- getBM(mart=grch37, attributes=c('ensembl_gene_id','go_id'))
  saveRDS(EG2GO, file='EG2GO.RDS')
}else{
  EG2GO <- readRDS('EG2GO.RDS')
}
```


## 4.2 Convert Ensembl Id into Go Id (or name)
```{r, eval=FALSE}
EG2GO <- EG2GO[EG2GO$go_id != '',]

upGoId <- EG2GO %>%
  filter(ensembl_gene_id %in% UpGeneList) %>%
  dplyr::select(go_id)
upGoId <- unique(upGoId)

downGoId <- EG2GO %>%
  filter(ensembl_gene_id %in% DownGeneList) %>%
  dplyr::select(go_id)
downGoId <- unique(downGoId)

allGoId <- EG2GO %>%
  filter(ensembl_gene_id %in% AllGeneList) %>%
  dplyr::select(go_id)
allGoId <- unique(allGoId)

```



## 4.3 Using topGo package to do enrichment tests
```{r, eval=FALSE}
xx <- annFUN.org("BP", mapping = "org.Hs.eg.db", ID = "symbol")
head(xx)

geneList <- factor(as.integer( allGoId[,1] %in% upGoId[,1]))
names(geneList) <- allGoId

str(geneList)

upGoData <- new("topGOdata",
                 description='GO data for increased genes',
                 ontology = "BP", allGenes = geneList, geneSel = upGoId, nodeSize = 5,
                 annot = annFUN.db, mapping = "org.Hs.eg.db", ID = "symbol")


resultFisher <- runTest(upGoData, algorithm = "classic", statistic = "fisher")
resultFisher

resultKS <- runTest(sampleGOdata, algorithm = "classic", statistic = "ks")
resultKS.elim <- runTest(sampleGOdata, algorithm = "elim", statistic = "ks")

showSigOfNodes(upGoData, score(resultKS.elim), firstSigNodes = 5, useInfo = 'all')

```

# 5. Using GOStat package to do enrichment tests

## 5.1 Map Ids

```{r}
res$symbol = mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")
res$entrez = mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")
res$goids = mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     column="GO",
                     keytype="ENSEMBL",
                     multiVals="first")
res$name =   mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     column="GENENAME",
                     keytype="ENSEMBL",
                     multiVals="first")

```

## 5.2 Up regulated list and Down regulated list

Up: filter by P.adj < 0.05 and LFC > 1  
Down: filter by P.adj < 0.05 and LFC < -1  

```{r}
up.idx <- which(res$padj<0.05 & res$log2FoldChange > 1)
up.entrez.id <- unique(res[up.idx, 'entrez'])

down.idx <- which(res$padj<0.05 & res$log2FoldChange < -1)
down.entrez.id <- unique(res[down.idx, 'entrez'])

universeGenes <-  unique(res$entrez)

```

## 5.3 Up gene GO analysis

### BP
```{r}
params <- new("GOHyperGParams",
     geneIds=up.entrez.id,
     universeGeneIds=universeGenes,
     annotation="org.Hs.eg.db",
     ontology="BP",
     pvalueCutoff=0.001,
     conditional=FALSE,
     testDirection="over")

hgOver <- hyperGTest(params)
summary(hgOver)[1:10,c("GOBPID","Pvalue","Term")]
```

### CC
```{r}
params <- new("GOHyperGParams",
     geneIds=up.entrez.id,
     universeGeneIds=universeGenes,
     annotation="org.Hs.eg.db",
     ontology="CC",
     pvalueCutoff=0.001,
     conditional=FALSE,
     testDirection="over")

hgOver <- hyperGTest(params)
summary(hgOver)[1:10,c("GOCCID","Pvalue","Term")]
```

### MF
```{r}
params <- new("GOHyperGParams",
     geneIds=up.entrez.id,
     universeGeneIds=universeGenes,
     annotation="org.Hs.eg.db",
     ontology="MF",
     pvalueCutoff=0.001,
     conditional=FALSE,
     testDirection="over")

hgOver <- hyperGTest(params)
summary(hgOver)[1:10,c("GOMFID","Pvalue","Term")]
```

## 5.4 Down gene GO analysis

### BP
```{r}
params <- new("GOHyperGParams",
     geneIds=down.entrez.id,
     universeGeneIds=universeGenes,
     annotation="org.Hs.eg.db",
     ontology="BP",
     pvalueCutoff=0.001,
     conditional=FALSE,
     testDirection="over")

hgOver <- hyperGTest(params)
summary(hgOver)[1:10,c("GOBPID","Pvalue","Term")]
```

### CC
```{r}
params <- new("GOHyperGParams",
     geneIds=down.entrez.id,
     universeGeneIds=universeGenes,
     annotation="org.Hs.eg.db",
     ontology="CC",
     pvalueCutoff=0.001,
     conditional=FALSE,
     testDirection="over")

hgOver <- hyperGTest(params)
summary(hgOver)[1:10,c("GOCCID","Pvalue","Term")]
```

### MF
```{r}
params <- new("GOHyperGParams",
     geneIds=down.entrez.id,
     universeGeneIds=universeGenes,
     annotation="org.Hs.eg.db",
     ontology="MF",
     pvalueCutoff=0.001,
     conditional=FALSE,
     testDirection="over")

hgOver <- hyperGTest(params)
summary(hgOver)[1:10,c("GOMFID","Pvalue","Term")]
```





# 6. Pathway: Gage



## 6.1 Gage anaylsis using kegg

```{r}
library(gageData)
resLFC = res$log2FoldChange
names(resLFC) = res$entrez
data(kegg.sets.hs)

kres = gage(resLFC, gsets=kegg.sets.hs, same.dir=TRUE)


k_pathways_Up = data.frame(id=rownames(kres$greater), kres$greater) %>% 
  tbl_df() %>% 
  filter(row_number()<=5) %>% 
  .$id %>% 
  as.character()

k_pathways_Down = data.frame(id=rownames(kres$less), kres$less) %>% 
  tbl_df() %>% 
  filter(row_number()<=5) %>% 
  .$id %>% 
  as.character()

k_Up_ids = substr(k_pathways_Up, start=1, stop=8)
k_Down_ids = substr(k_pathways_Down, start=1, stop=8)

```


### Up regulated pathways

```{r}
pathview(gene.data=resLFC, pathway.id=k_Up_ids, species="hsa", kegg.dir = "PathviewResult/Up/", new.signature=FALSE)
for(f in k_Up_ids){
  f <- paste0(f,'.pathview.png')
  knitr::include_graphics(f)
}

knitr::include_graphics(f)
```


### Down regulated pathways

```{r}
pathview(gene.data=resLFC, pathway.id=k_Down_ids, species="hsa", kegg.dir = "PathviewResult/Down/", new.signature=FALSE)
for(f in k_Down_ids){
  f <- paste0(f,'.pathview.png')
  img <- readPNG(f)
  plot.new()
  rasterImage(img,-0.5, -0.3, 1.5, 1.3)
}

```

## 6.2 Gage anaylsis using GO
```{r}
data(go.sets.hs)
data(go.subs.hs)
resLFC = res$log2FoldChange

gobpsets = go.sets.hs[go.subs.hs$BP]
goccsets = go.sets.hs[go.subs.hs$CC]
gomfsets = go.sets.hs[go.subs.hs$MF]

GoLFC <- resLFC
names(GoLFC) <- res$goids

gobpres <- gage(GoLFC, gsets=gobpsets, same.dir=TRUE)
goccres = gage(GoLFC, gsets=goccsets, same.dir=TRUE)
gomfres = gage(GoLFC, gsets=gomfsets, same.dir=TRUE)


gobpres = gage(resLFC, gsets=gobpsets, same.dir=TRUE)

```

### bp result

```{r}

lapply(gobpres, head)
```

### cc result
```{r}
lapply(goccres, head)
```

### mf result

```{r}
lapply(gomfres, head)
```

